---
title: "Tidying of Job Patterns Data - MA615 Midterm Project"
author: "Munan Hou"
date: "October 16, 2016"
output: html_document
---

#Information of the dataset
Topic: "Job Patterns For Minorities And Women In Private Industry, 2009 EEO-1 NAICS-4 Aggregate Report"
File name: YEAR09_CBSA_NAC3.txt (with data_dictionary NAICS.xls)
Resource from: 
https://catalog.data.gov/dataset/job-patterns-for-minorities-and-women-in-private-industry-2009-eeo-1-naics-4-aggregate-rep

#Import Data
First of all we import the data from the .txt file, using read.table here. It is noticed that the separation of this dataset is ";" instead of ",", so we added the "sep";" argument. There is no missing values inside the dataset, or the missing values are valued instead of leaving blank.

```{r import txt}
library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2)
data <- read.table("data/YEAR09_CBSA_NAC3.txt", header = TRUE, sep = ";", quote = "", dec = ".", fill = TRUE)
data_backup <- data
```

This dataframe contains 600 columns(variables), 9064 rows(observations).

#Analysis of the dataset

In my understianding, because the dataset are named with "minorities and women", this dataset should be generally used to analyse:

- the probable different job patterns among races, under each industry, or under each job category.
- the probable different job patterns between genders, under each industry, or under each job category.
- the probable different job patterns under the two kinds of previous precondition, and further under different geographic positions.

```{r}
head(data, n=1)
```

By looking at the data dictionary NAICS.xls, and also by reading the dataset, we can have the information below:

Three categories under gender: Male, Female and Total

The industries are separated into 11 different categories within each race and gender, labeled as *1, *2, ... *10, *1_2 (* represents the wildcard character). Same number represent the same categoreis although in different industries.

As long as our purpose of tidying this data is known, the tidying steps are as follow. After data being tidied up, the data should become more clear to model, compare and test for difference's significance.

#First step, analyse the 600 columns

The first obvious mess thing is that, it is sure not each of columns should be a single variable. Inside columns, there are both variables and observations. Within them, "CBSA_LABEL" are area information, "NAC3_Label" are industry information; they are for sure fixed variables(dimensions) of the data. For variables "X_TYPE_", "TOTAL_UNIT", "FIRMS", "pctofst", I can't figure out what do they mean; however,

For "X_TYPE_"

```{r}
count(data, X_TYPE_)
```

there is only one kind of outcome. So I think it could be the input type of the observation such as "integer", etc., os it is proper to treat it as a fixed variable and leave it unchanged. 

For "TOTAL_UNIT", 

```{r}
count(data, TOTAL_UNIT)
min(data$TOTAL_UNIT)
max(data$TOTAL_UNIT)
```

For "FIRMS"
"
```{r}
count(data, FIRMS)
min(data$FIRMS)
max(data$FIRMS)
```

For "pctofst", could be "percent of state"

```{r}
min(data$pctofst)
max(data$pctofst)
sum(data$pctofst)
```

No conclusion found. Leave unchanged. While as long as it seems could be a proper variable(although not fixed), I'll leave and regard it as already tidied.

Then for the rest of the listed variables. Three different variables should be extracted: 

- RACE("TOTAL", "WH(White)", "BLK(Black)", "HISP(Hispanic)", "ASIAN", "AIAN(Indian)", "nhopi(Hawaiian)", "tomr(Two Race)", "Min(Minority)", "WHp(%)", "BLKp(%)", "HISPp(%)", "AIANp(%)", "nhopip(%)", "asianp(%)", "tomrp(%)", "Minp(%)", "PTOTAL(%)"), 
- GENDER("F", "M", "T = F + M"), 
- INDUSTRY("*1, *2, ... *10, *1_2"s)

In my first thought, there should be actually these fixed and measured variable column:

- CBSA_LABEL
- NAC3_Label
- X_TYPE_

- FIRMS
- pctofst

- TOTAL_UNIT
- INDUSTRY
- GENDER
- RACE

- count

So I'll try to start tidy now.

First, I added one column called 'label'. I'll merge the separated dataframe by this identical variable.

```{r mark label}
data$label <- c(1:nrow(data)) #mark for further merge
data <- data[,c(601,1:600)] #rearrange the label vector to the first
```

But I notice that the sequence of each "total", "male" and "female" are not in order. For further convenience when looping, I make up the columns first, letting them in a tidy order of "T", "M", "F".

```{r}
name_list <- data.frame(number=c(1:(601)), variable.names(data)[1:601])
head(name_list)
```

Then read manully and rearrange.

```{r}
data <- data[,c(1:280, 292:302, 281:291,
                
                479,481:489,480, #WHtp
                303,305:313,304, #WHMp
                391,393:401,392, #WHfp
                
                (479+11*1),(481+11*1):(489+11*1),(480+11*1), #BLKtp
                (303+11*1),(305+11*1):(313+11*1),(304+11*1), #BLKMp
                (391+11*1),(393+11*1):(401+11*1),(392+11*1), #BLKfp
                
                (479+11*2),(481+11*2):(489+11*2),(480+11*2), #HISPtp
                (303+11*2),(305+11*2):(313+11*2),(304+11*2), #HISPMp
                (391+11*2),(393+11*2):(401+11*2),(392+11*2), #HISPfp
                
                (479+11*3),(481+11*3):(489+11*3),(480+11*3), #AIANtp1
                (303+11*3),(305+11*3):(313+11*3),(304+11*3), #AIANMp1
                (391+11*3),(393+11*3):(401+11*3),(392+11*3), #AIANfp1
                
                (479+11*4),(481+11*4):(489+11*4),(480+11*4), #nhopitp
                (303+11*4),(305+11*4):(313+11*4),(304+11*4), #nhopiMp
                (391+11*4),(393+11*4):(401+11*4),(392+11*4), #nhopifp
                
                (479+11*5),(481+11*5):(489+11*5),(480+11*5), #asiantp
                (303+11*5),(305+11*5):(313+11*5),(304+11*5), #asianMp
                (391+11*5),(393+11*5):(401+11*5),(392+11*5), #asianfp
                
                (479+11*6),(481+11*6):(489+11*6),(480+11*6), #tomrtp
                (303+11*6),(305+11*6):(313+11*6),(304+11*6), #tomrMp
                (391+11*6),(393+11*6):(401+11*6),(392+11*6), #tomrfp
                
                (479+11*7),(481+11*7):(489+11*7),(480+11*7), #MinTp
                (303+11*7),(305+11*7):(313+11*7),(304+11*7), #MinMp
                (391+11*7),(393+11*7):(401+11*7),(392+11*7), #MinFp
                
                589,591:599,590, #PTOTAL
                567,569:577,568, #MTp
                578,580:588,579,  #FTp
                
                600,601
                )]
```

There is order in it, so not too difficult to do this.

```{r}
name_list_test <- data.frame(number=c(1:(601)), variable.names(data)[1:601])
name_list_test
```

Succeeded.

Then I define 

```{r}
VAR_NAME_ALL <- c(variable.names(data))
VAR_NAME_temp <- VAR_NAME_ALL[seq(6,589,11)]
VAR_NAME_temp
```

This "VAR_NAME_temp will" and the VAR_NAME and RACE_NAME be used later, in order to rename the merged variables.

So I have extracted names to be corrected manully. 

```{r}
VAR_NAME <- c("TOTAL","MT","FT","WHT","WHM","WHF","BLKT","BLKM","BLKF","HISPT","HISPM","HISPF","ASIANT","ASIANM","ASIANF","AIANT","AIANM","AIANF","nhopiT","NHOPIM","NHOPIF","tomrT","TOMRM","TOMRF","MinT","MinM","MinF","WHtp","WHMp","WHfp","BLKtp","BLKMp","BLKfp","HISPtp","HISPMp","HISPfp","AIANtp","AIANMp","AIANfp","nhopitp","nhopiMp","nhopifp","asiantp","asianMp","asianfp","tomrtp","tomrMp","tomrfp","MinTp","MinMp","MinFp","PTOTAL","MTp","FTp")
RACE_NAME <- c("TOTAL","WH","BLK","HISP","ASIAN","AIAN","nhopi","tomr","Min","WHp","BLKp","HISPp","AIANp","nhopip","asianp","tomrp","Minp","PTOTAL")
```

#Trying to tidy all variables up

Now tidying.

As I have analyzed before, at first I believe the dataset should be totally tidied, as all the 11 different "INDUSTRY" categories within each race and gender are spreading in columns, labeled as *1, *2, ... *10, *1_2, these should be merged into variables of "INDUSTRY" and "count", just as what need to be done on "RACE" and "GENDER"
I first splitted all different "RACE"s and "GENDER"s into separated spreadsheets, define the first chain of the loop by merging its "*1, *2, ... *10, *1_2" into "INDUSTRY" and "count", then use if loop and similar method to automatically merge all other "INDUSTRY"s and "count"s among each combination of "RACE" and "GENDER". At the same time, I add "RACE" and "GENDER" columns into each spreadsheets, for future combine into two overall "RACE" and "GENDER" variable.

#Method 1

```{r first step tidying TOTALLY everything}
data_TOTAL_T <- data[c(1, 6:16)]
data_TOTAL_M <- data[c(1, 17:27)]
data_TOTAL_F <- data[c(1, 28:38)]

INDUSTRY_NAME <- c("label", 1:11) #I first rename the column names, then merge them, then rename again
names(data_TOTAL_T) <- INDUSTRY_NAME
names(data_TOTAL_M) <- INDUSTRY_NAME
names(data_TOTAL_F) <- INDUSTRY_NAME

data_TOTAL_T$GENDER <- "T"
data_TOTAL_M$GENDER <- "M"
data_TOTAL_F$GENDER <- "F"
data_TOTAL_T$RACE <- "TOTAL"
data_TOTAL_F$RACE <- "TOTAL"
data_TOTAL_M$RACE <- "TOTAL"

data_TOTAL_T <- gather(data_TOTAL_T, INDUSTRY, count, 2:12)
data_TOTAL_F <- gather(data_TOTAL_F, INDUSTRY, count, 2:12)
data_TOTAL_M <- gather(data_TOTAL_M, INDUSTRY, count, 2:12)

data_TOTAL <- bind_rows(data_TOTAL_T, data_TOTAL_F, data_TOTAL_M)
head(data_TOTAL)
```

```{r loop of tidying everything}
for (i in 1:17) {
data_TOTAL_T <- data[c(1, (33*i + 6):(33*i + 16))]
data_TOTAL_M <- data[c(1, (33*i + 17):(33*i + 27))]
data_TOTAL_F <- data[c(1, (33*i + 28):(33*i + 38))]

names(data_TOTAL_T) <- INDUSTRY_NAME
names(data_TOTAL_M) <- INDUSTRY_NAME
names(data_TOTAL_F) <- INDUSTRY_NAME

data_TOTAL_T$GENDER <- "T"
data_TOTAL_M$GENDER <- "M"
data_TOTAL_F$GENDER <- "F"
data_TOTAL_T$RACE <- RACE_NAME[i + 1]
data_TOTAL_F$RACE <- RACE_NAME[i + 1]
data_TOTAL_M$RACE <- RACE_NAME[i + 1]

data_TOTAL_T <- gather(data_TOTAL_T, INDUSTRY, count, 2:12)
data_TOTAL_F <- gather(data_TOTAL_F, INDUSTRY, count, 2:12)
data_TOTAL_M <- gather(data_TOTAL_M, INDUSTRY, count, 2:12)

data_TOTAL <- bind_rows(data_TOTAL, data_TOTAL_T, data_TOTAL_F, data_TOTAL_M)
  
}
head(data_TOTAL)
```

delete rows of count = 0 or NA

```{r}
data_TOTAL <- filter(data_TOTAL, is.na(count) == FALSE)
data_TOTAL <- filter(data_TOTAL, count != 0)
```

rename "INDUSTRY"

```{r rename INDUSTRY}
data_TOTAL$INDUSTRY[which(data_TOTAL$INDUSTRY == "1")] = "SENIOR OFF AND MGRS"
data_TOTAL$INDUSTRY[which(data_TOTAL$INDUSTRY == "2")] = "PROFESSIONALS"
data_TOTAL$INDUSTRY[which(data_TOTAL$INDUSTRY == "3")] = "TECHNICIANS"
data_TOTAL$INDUSTRY[which(data_TOTAL$INDUSTRY == "4")] = "SALES WORKERS"
data_TOTAL$INDUSTRY[which(data_TOTAL$INDUSTRY == "5")] = "OFFICE AND CLERICALS"
data_TOTAL$INDUSTRY[which(data_TOTAL$INDUSTRY == "6")] = "CRAFT WORKERS"
data_TOTAL$INDUSTRY[which(data_TOTAL$INDUSTRY == "7")] = "OPERATIVES"
data_TOTAL$INDUSTRY[which(data_TOTAL$INDUSTRY == "8")] = "LABORERS"
data_TOTAL$INDUSTRY[which(data_TOTAL$INDUSTRY == "9")] = "SERVICE WORKERS"
data_TOTAL$INDUSTRY[which(data_TOTAL$INDUSTRY == "10")] = "TOTAL"
data_TOTAL$INDUSTRY[which(data_TOTAL$INDUSTRY == "11")] = "MID OFF AND MGRS"
head(data_TOTAL)
```

Creat a "dictionary" that also marked by "label", then merge it with the dataset.

```{r make dictionary}
data_dict <- data[c(1:5,(ncol(data)-1),ncol(data))]
head(data_dict)
```

merge with dataframe "dictionary" by variable "label",

```{r merge}
data_TOTAL <- right_join(data_dict, data_TOTAL, by="label")
head(data_TOTAL)
```

So that's it.

But this could be too long. Also, it might be unnecessary to tidy industries because we don't need to test the difference among different industry.

#Method 2

For compare different races & genders in same industries, maybe only tidy the part of industries can be enough:

```{r first step before loop}
INDUSTRY_NAME <- c("label", 1:11)
data_sort_in <- data.frame(label = NA, INDUSTRY = NA)

data_sort_in_1 <- data[c(1, (6):(16))]
data_sort_in_2 <- data[c(1, (17):(27))]
data_sort_in_3 <- data[c(1, (28):(38))]

names(data_sort_in_1) <- names(data_sort_in_2) <- names(data_sort_in_3) <- INDUSTRY_NAME

data_sort_in_1 <- gather(data_sort_in_1, INDUSTRY, a, 2:12)
data_sort_in_2 <- gather(data_sort_in_2, INDUSTRY, b, 2:12)
data_sort_in_3 <- gather(data_sort_in_3, INDUSTRY, c, 2:12)

names(data_sort_in_1) <- c("label", "INDUSTRY", VAR_NAME[1])
names(data_sort_in_2) <- c("label", "INDUSTRY", VAR_NAME[2])
names(data_sort_in_3) <- c("label", "INDUSTRY", VAR_NAME[3])

data_sort_in <- merge(data_sort_in_1, data_sort_in_2, by=c("label", "INDUSTRY"))
data_sort_in <- merge(data_sort_in, data_sort_in_3, by=c("label", "INDUSTRY"))
```

```{r second step loop}

for (i in 1:17) {

data_sort_in_1 <- data[c(1, (33*i + 6):(33*i + 16))]
data_sort_in_2 <- data[c(1, (33*i + 17):(33*i + 27))]
data_sort_in_3 <- data[c(1, (33*i + 28):(33*i + 38))]

names(data_sort_in_1) <- names(data_sort_in_2) <- names(data_sort_in_3) <- INDUSTRY_NAME

data_sort_in_1 <- gather(data_sort_in_1, INDUSTRY, a, 2:12)
data_sort_in_2 <- gather(data_sort_in_2, INDUSTRY, b, 2:12)
data_sort_in_3 <- gather(data_sort_in_3, INDUSTRY, c, 2:12)

names(data_sort_in_1) <- c("label", "INDUSTRY", VAR_NAME[3*i + 1])
names(data_sort_in_2) <- c("label", "INDUSTRY", VAR_NAME[3*i + 2])
names(data_sort_in_3) <- c("label", "INDUSTRY", VAR_NAME[3*i + 3])

data_sort_in <- merge(data_sort_in, data_sort_in_1, by=c("label", "INDUSTRY"))
data_sort_in <- merge(data_sort_in, data_sort_in_2, by=c("label", "INDUSTRY"))
data_sort_in <- merge(data_sort_in, data_sort_in_3, by=c("label", "INDUSTRY"))

}
head(data_sort_in)
```

merge with dictionary

```{r merge with dictionary}
data_sort_in_final <- right_join(data_dict, data_sort_in, by="label")
head(data_sort_in_final)
```

```{r rename INDUSTRY}
data_sort_in_final$INDUSTRY[which(data_sort_in_final$INDUSTRY == "1")] = "SENIOR OFF AND MGRS"
data_sort_in_final$INDUSTRY[which(data_sort_in_final$INDUSTRY == "2")] = "PROFESSIONALS"
data_sort_in_final$INDUSTRY[which(data_sort_in_final$INDUSTRY == "3")] = "TECHNICIANS"
data_sort_in_final$INDUSTRY[which(data_sort_in_final$INDUSTRY == "4")] = "SALES WORKERS"
data_sort_in_final$INDUSTRY[which(data_sort_in_final$INDUSTRY == "5")] = "OFFICE AND CLERICALS"
data_sort_in_final$INDUSTRY[which(data_sort_in_final$INDUSTRY == "6")] = "CRAFT WORKERS"
data_sort_in_final$INDUSTRY[which(data_sort_in_final$INDUSTRY == "7")] = "OPERATIVES"
data_sort_in_final$INDUSTRY[which(data_sort_in_final$INDUSTRY == "8")] = "LABORERS"
data_sort_in_final$INDUSTRY[which(data_sort_in_final$INDUSTRY == "9")] = "SERVICE WORKERS"
data_sort_in_final$INDUSTRY[which(data_sort_in_final$INDUSTRY == "10")] = "TOTAL"
data_sort_in_final$INDUSTRY[which(data_sort_in_final$INDUSTRY == "11")] = "MID OFF AND MGRS"
head(data_sort_in_final)
```

re-arrange

```{r re-arrange variables, remove label}
data_sort_in_final <- data_sort_in_final[,c(2:5),c(8:ncol(data_sort_in_final),c(6:7))]
head(data_sort_in_final)
```

So this is the final tidy dataset in this situation. But in this form, we cannot easily remove all the 0's and NA's. So compare to Method 1, Method 2 is not as good as the previous method.

#Result

Choose Method 1

#Visualization

For data_TOTAL,








